\documentclass[10pt,twocolumn,letterpaper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=2cm,right=2cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{float}

%% Title
\title{
        %\vspace{-1in}     
        \usefont{OT1}{bch}{b}{n}
        \normalfont \normalsize \textsc{Advanced Computer Graphics Spring 2019 (RPI)} \\ [10pt]
        \huge Physics based Character Posing and Inverse Kinematics: Proposal \\
}

\usepackage{authblk}
\author[0]{Frederick Choi}


\begin{document}
\maketitle

\selectlanguage{english}

\section*{Introduction}
Posing and animating rigs can be a pain for animators. Inverse kinematics solvers calculate the relative rotations of rigid bodies connected by different types of joints, which is useful especially for robotics and animation. However, when posing a rig for modeling or animation, some inverse kinematics solvers can produce unpredictable, and unintuitive, especially for large displacements. Inspired by how artists manually pose wooden armatures, I propose a physics-based character posing and inverse kinematics algorithm.

Cooper and Ballard \cite{Co} published a marker-based following algorithm that similarly constructs a physical model to simulate markers dragging the rig into position and use the Open Dynamic Physics engine to run their simulation. My algorithm will expand on this paper by focusing on posing. As described below, this project will have an implementation of a simple physics engine with modifications that ensure stability and will interpolate the motion of the control point to ensure predictable convergence.

iTasC published by Smith et al. \cite{Sm} includes an algorithm for solving inverse kinematics, which is available in Blender. I will be comparing the performance of my algorithm with the iTasC algorithm in Blender based on the objective points below. Kulpa and Multon \cite{Ku} have a hierarchical algorithm for IK on humanoid rigs that is both quick but also calculates realistic poses for humans. FABRIK by Artistidou and Lasenby \cite{Ar} have an inverse-kinematics solver that is more geared towards robotics. These are both iterative methods that approximate the solution by moving each piece "into place" then correcting the other parts of the rig. All of these are useful both as a baseline as they are fast algorithms that produce stable solutions, whereas a physics based solutions might not even converge. Also, to improve the performance of the algorithm in this paper, a hybrid of physics and the above techniques might be used.

Pickl's master thesis \cite{Pi} has a lot of formulas describing the dynamics of rigid bodies connected by links and joints. Since the algorithm will be simulating rigs, which are basically just rigid bodies connected by joints, that paper will be the primary reference for rigid body dynamics.

\section*{Objectives} The objectives of the physics-based character posing algorithm are:
\begin{itemize}
    \item Accuracy. The points on the rig should end up close to their target position while satisfying all constraints.
    \item Predictability. The solution should be close to the user's expectations (e.g no jumpy motions, things turning inside out).
    \item Performance. The algorithm should not be prohibitively slow.
\end{itemize}

Accuracy can be evaluated quantitatively. Predictability will be evaluated qualitatively by comparing it with other IK solvers in Blender. While speed is also a consideration, it will not be the focus of this project, since the limitations of the Blender plugin API makes it difficult to measure the true performance of the algorithm.

\section*{Algorithm}

\paragraph{Interface.} The posing tool will be implemented as a Blender plugin. The user will be able to place control points on the rig and move them around. 

\paragraph{Control Points.} A control point consists of two parts: the attachment point and the target point. The attachment point is on the rig, while the target point is what the user moves around. The attachment point is brought toward the target point by simulating a spring (linear, constant, or non-linear). Multiple control points will be supported.

\paragraph{Movement Interpolation.} When the target point of a control point is moved far away from a control point (i.e. the angular displacement exceeds some threshold) then the movement of the target point will be interpolated over time to ensure the algorithm converges to a predictable solution. Some ideas to experiment with include:
\begin{itemize}
    \item A circular arc around the center of mass
    \item A circular arc around the joint
    \item Linear time interpolation
    \item Ease-in / ease-out time interpolation
\end{itemize}

\paragraph{Physics.} A simple, custom physics engine will be implemented as well so that it can be modified to ensure stability, convergence, and performance. The accuracy of the physics in between is only important insofar as it affects the outcome. The mechanics that will be modelled are:
\begin{itemize}
    \item External torques
    \item Inter-joint forces
    \item Torsion forces
\end{itemize}
Modelling masses, stiffness, etc. will ensure that things will not move unneseccesarily. For example, the whole body will not move just to pose a finger. This improves the predictability of the computed pose.

\paragraph{Iterative Methods.} After running a physical simulation to reach an approximate, predictable configuration, iterative methods (as described in \cite{Sm}\cite{Ku}\cite{Ar}) will be used to get an exact solution, if possible. Otherwise, if the control points are placed in a configuration inconsistent with the users constraints, the physical simulation should still produce predictable results.

\section*{Project Timeline}

The following is an outline of what I will accomplish each week.

\paragraph{Week 0.} I already worked through the Blender API and I have a simple plugin working. I also currently have things moving, and multi-joint physics working on the plane.

\paragraph{Week 1.} Get multi-joint physics working in 3D space and get axial torques to work. Model rotation constraints as counter-rotation springs. To demonstrate, I will pose a simple two-joint model (in real time). I will also get physics working on branching/multi joints (hip/shoulders). To demonstrate, I will pose a simple field-goal shaped model.

\paragraph{Week 2.} Ensure mass annd moment of inertia works properly. Implement stiffness. Work on interpolation of marker movement and play around with different types of springs. To demonstrate, I will have comparisons between several of the above alternatives.

\paragraph{Week 3.} Implement iterative algorithms to improve performance and get exact solutions whenever possible.

\paragraph{Presentation.} Live demo of posing a simple human skeleton (limbs, torso, no fingers).

\bibliography{research}
\bibliographystyle{ieeetr}

\end{document}